Capstone Movielens Report

### 2025-12-13 reports/capstone-movielens.site/05.appendix.Rmd
Function described: `train_UMGY_SmoothedDay_effect.RMSE.cv`.

>>>>

#### [train_UMGY_SmoothedDay_effect.RMSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L173) Function {#func.train_UMGY_SmoothedDay_effect.RMSE.cv}
\

Trains and validate (without *regularization*) the *User+Movie+Genre+Year+SmoothedDay Effect (UMGYDE) Model* for the given parameter values passed in the [degree](#) and [span](#) arguments, with use of *K-Fold Cross-Validation* as described in [Section 29.6 *Cross validation / K-fold cross validation*](https://rafalab.dfci.harvard.edu/dsbook-part-2/ml/resampling-methods.html#k-fold-cross-validation) of the _Course Textbook (New Edition)_[@IDS2_29-6_KFCV].

::: { #func.train_UMGY_SmoothedDay_effect.RMSE.cv.usage .sidebar }
##### Usage
\

```{r eval=FALSE}
  train_UMGY_SmoothedDay_effect.RMSE.cv(degree = 1, span)
```

::: {.noteblock data-latex=""}
In the above code snippet, the `train_UMGY_SmoothedDay_effect.RMSE.cv` function is called internally from the [train_UMGY_SmoothedDay_effect.RMSE.cv.degree1](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.degree1) function described below in this section.
:::
:::

::: { #func.train_UMGY_SmoothedDay_effect.RMSE.cv.args .sidebar }
##### Arguments
\

::: { #func.train_UMGY_SmoothedDay_effect.RMSE.cv.args.degree .sidebar }
  - **degree:** (Integer, $NA$ by default) The degree of the polynomials to pass in the same-named argument to the internally called [train_UMGY_SmoothedDay_effect.cv](#func.train_UMGY_SmoothedDay_effect.cv) described above;

:::
::: { #func.train_UMGY_SmoothedDay_effect.RMSE.cv.args.span .sidebar }
  - **span:** (Numeric, $NA$ by default ) The parameter $\alpha$ (which controls the degree of smoothing) to pass in the same-named argument to the internally called [train_UMGY_SmoothedDay_effect.cv](#func.train_UMGY_SmoothedDay_effect.cv) described above.

::: { #func.train_UMGY_SmoothedDay_effect.RMSE.cv.details .sidebar }
##### Details
\

This is a wrapper for the function [train_UMGY_SmoothedDay_effect.cv](#func.train_UMGY_SmoothedDay_effect.cv) (described above in section [UMGYDE Model: Utility Functions] of *this Appendix*), which it calls internally, passing the values of the arguments [degree](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.args.degree) and [span](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.args.span) in the same-named arguments.

Unlike the internal function  [train_UMGY_SmoothedDay_effect.cv](#func.train_UMGY_SmoothedDay_effect.cv), this one also validates the *UMGYDE Model* created and trained by that function, for which it calls internally another helper function, [calc_UMGY_SmoothedDay_effect.RMSE.cv](#func.calc_UMGY_SmoothedDay_effect.RMSE.cv) (also described above in the [UMGYDE Model: Utility Functions] section), computing *RMSE* for that model.

::: {.noteblock data-latex=""}
This is an auxiliary  function intended to be called internally from the following helper functions (described below in this section), specifically designed for tuning the `degree` and `span` [loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess) parameters:

  - [train_UMGY_SmoothedDay_effect.RMSE.cv.degree0](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.degree0): a helper function designed for tuning the [loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess) `span` parameter while keeping the `degree` parameter fixed at $0$;

  - [train_UMGY_SmoothedDay_effect.RMSE.cv.degree1](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.degree1): a helper function designed for tuning the [loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess) `span` parameter while keeping the `degree` parameter fixed at $1$;

  - [train_UMGY_SmoothedDay_effect.RMSE.cv.degree2](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.degree2): a helper function designed for tuning the [loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess) `span` parameter while keeping the `degree` parameter fixed at $2$;

::: { #func.train_UMGY_SmoothedDay_effect.RMSE.cv.value .sidebar }
##### Value
\

The *RMSE* value calculated for the *UMGYDE Model* fitted for the given [loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess) parameters passed in the [degree](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.args.degree) and [span](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.args.span) arguments, and trained with use of *K-Fold Cross-Validation*, where the $K$ is the length of the [`edx_CV` Object] (in *this Project* we use $K = 5$).
:::

::: { #func.train_UMGY_SmoothedDay_effect.RMSE.cv.code .sidebar }
##### Source Code
\

The slightly simplified version of the source code of the [train_UMGY_SmoothedDay_effect.RMSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L173) function is shown below:
```{r eval=FALSE}
train_UMGY_SmoothedDay_effect.RMSE.cv <- function(degree = NA, span = NA) {
  smth_de.RMSE <- train_UMGY_SmoothedDay_effect.cv(degree, span) |>
    calc_UMGY_SmoothedDay_effect.RMSE.cv()

  put_log2("Function `train_UMGY_SmoothedDay_effect.RMSE.cv`:
RMSE has been computed for the `loess` function parameters: 
degree = %1, span = %2.", 
           degree,
           span)

  smth_de.RMSE
}
```

::: {.noteblock data-latex=""}
The complete source code of the [train_UMGY_SmoothedDay_effect.RMSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L173) function is available in the [Utility Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L3) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L173) script on _GitHub_.
:::
:::
